---
title: "GlareosaeBiogeography"
author: "amesclir"
date: "03/02/2016"
output: html_document
---

First, load the packages.

```{r}
library(optimx)
library(FD)
library(snow)
library(parallel)
library(BioGeoBEARS)

source("http://phylo.wdfiles.com/local--files/biogeobears/cladoRcpp.R") 

source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_add_fossils_randomly_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_basics_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_calc_transition_matrices_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_classes_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_detection_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_DNA_cladogenesis_sim_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_extract_Qmat_COOmat_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_generics_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_models_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_on_multiple_trees_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_plots_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_readwrite_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_simulate_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_SSEsim_makePlots_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_SSEsim_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_stochastic_mapping_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_stratified_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_univ_model_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/calc_uppass_probs_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/calc_loglike_sp_v01.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/get_stratified_subbranch_top_downpass_likelihoods_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/runBSM_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/stochastic_map_given_inputs.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/summarize_BSM_tables_v1.R")
calc_loglike_sp = compiler::cmpfun(calc_loglike_sp_prebyte)    # crucial to fix bug in uppass calculations
calc_independent_likelihoods_on_each_branch = compiler::cmpfun(calc_independent_likelihoods_on_each_branch_prebyte)
    # slight speedup hopefully

```

First, load the tree and distribution.

```{r}

tr = read.nexus("tree.tree")
tr
tipstoremove <- c("Cremota","Cspicata", "Cpaniculata", "Celongata", "Cscoparia", "Cmerritt-fernaldi", "Cbolanderi", "Cleptopoda", "Cdeweyana", "Cbromoides", "Cchordorrhyza")
tr = drop.tip (tr, tipstoremove)
plot(tr, cex = 0.8)
title("Example Glareosae phylogeny")
axisPhylo()

write.tree(tr, file = "glareosae.tree")

tr = read.tree("glareosae.tree")
tr
plot(tr, cex = 0.8)
title("Example Glareosae phylogeny")
axisPhylo()

trfn = np("glareosae.tree")
moref(trfn)

geogfn = np("glareosae.data")
#extdata_dir = np(system.file("extdata", package="BioGeoBEARS"))
#extdata_dir
#list.files(extdata_dir)
#geogfn = np(paste(addslash(extdata_dir), "Psychotria_geog.data", sep=""))

geogfn
moref(geogfn)

# Look at your geographic range data:
tipranges = getranges_from_LagrangePHYLIP(lgdata_fn=geogfn)
tipranges

max_range_size = 5

numstates_from_numareas(numareas=5, maxareas=5, include_null_range=TRUE)
numstates_from_numareas(numareas=5, maxareas=5, include_null_range=FALSE)


```

3. Ahora vamos a usar el modelo DEC

```{r}

#######################################################
# Run DEC
#######################################################

# Intitialize a default model (DEC model)
BioGeoBEARS_run_object = define_BioGeoBEARS_run()

# Give BioGeoBEARS the location of the phylogeny Newick file
BioGeoBEARS_run_object$trfn = trfn

# Give BioGeoBEARS the location of the geography text file
BioGeoBEARS_run_object$geogfn = geogfn

# Input the maximum range size
BioGeoBEARS_run_object$max_range_size = max_range_size

BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
# 1. Here, un-comment ONLY the files you want to use.
# 2. Also un-comment "BioGeoBEARS_run_object = section_the_tree(...", below.
# 3. For example files see (a) extdata_dir, 
#  or (b) http://phylo.wikidot.com/biogeobears#files
#  and BioGeoBEARS Google Group posts for further hints)
#
# Uncomment files you wish to use in time-stratified analyses:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
# (use more cores to speed it up; this requires
# library(parallel) and/or library(snow). The package "parallel" 
# is now default on Macs in R 3.0+, but apparently still 
# has to be typed on some Windows machines. Note: apparently 
# parallel works on Mac command-line R, but not R.app.
# BioGeoBEARS checks for this and resets to 1
# core with R.app)

# Sparse matrix exponentiation is an option for huge numbers of ranges/states (600+)
# I have experimented with sparse matrix exponentiation in EXPOKIT/rexpokit,
# but the results are imprecise and so I haven't explored it further.
# In a Bayesian analysis, it might work OK, but the ML point estimates are
# not identical.
# Also, I have not implemented all functions to work with force_sparse=TRUE.
# Volunteers are welcome to work on it!!
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC model
# (nothing to do; defaults)

# Look at the BioGeoBEARS_run_object; it's just a list of settings etc.
BioGeoBEARS_run_object

# This contains the model object
BioGeoBEARS_run_object$BioGeoBEARS_model_object

# This table contains the parameters of the model 
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table

# Run this to check inputs. Read the error messages if you get them!
check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "Glareosae_DEC_M0_unconstrained_v1.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDEC = res
    } else {
    # Loads to "res"
    load(resfn)
    resDEC = res
    }



```

4. Run DEC + J

```{r}



#######################################################
# Run DEC+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDEC$outputs@params_table["d","est"]
estart = resDEC$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Add j as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
resfn = "Glareosae_DEC+J_M0_unconstrained_v1.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDECj = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECj = res
    }
```


5. Plotting the results

```{r}

#######################################################
# PDF plots
#######################################################
pdffn = "Glareosae_DEC_vs_DEC+J_M0_unconstrained_v1.pdf"
pdf(pdffn, width=6, height=6)

#######################################################
# Plot ancestral states - DEC
#######################################################
analysis_titletxt ="BioGeoBEARS DEC on Glareosae M0_unconstrained"

# Setup
results_object = resDEC
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DECJ
#######################################################
analysis_titletxt ="BioGeoBEARS DEC+J on Glareosae M0_unconstrained"

# Setup
results_object = resDECj
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()  # Turn off PDF
cmdstr = paste("open ", pdffn, sep="")
system(cmdstr) # Plot it

```

6. Run DIVA like

```{r}
#######################################################
# Run DIVALIKE
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE model
# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# No jump dispersal/founder-event speciation
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = 0.01
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = 0.01

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

runslow = TRUE
resfn = "Glareosae_DIVALIKE_M0_unconstrained_v1.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDIVALIKE = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKE = res
    }

```

7. Run DIVALIKE + j

```{r}
#######################################################
# Run DIVALIKE+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDIVALIKE$outputs@params_table["d","est"]
estart = resDIVALIKE$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add jump dispersal/founder-event speciation
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Under DIVALIKE+J, the max of "j" should be 2, not 3 (as is default in DEC+J)
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","min"] = 0.00001
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","max"] = 1.99999

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

resfn = "Glareosae_DIVALIKE+J_M0_unconstrained_v1.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDIVALIKEj = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEj = res
    }


```

8. Plotting additional results

```{r}

pdffn = "Glareosae_DIVALIKE_vs_DIVALIKE+J_M0_unconstrained_v1.pdf"
pdf(pdffn, width=6, height=6)

#######################################################
# Plot ancestral states - DIVALIKE
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE on Glareosae M0_unconstrained"

# Setup
results_object = resDIVALIKE
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DIVALIKE+J
#######################################################
analysis_titletxt ="Glareosae DIVALIKE+J on Psychotria M0_unconstrained"

# Setup
results_object = resDIVALIKEj
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()
cmdstr = paste("open ", pdffn, sep="")
system(cmdstr)

```



9. Comparing all results

```{r}
#########################################################################
#########################################################################
#########################################################################
#########################################################################
# 
# CALCULATE SUMMARY STATISTICS TO COMPARE
# DEC, DEC+J, DIVALIKE, DIVALIKE+J
# 
#########################################################################
#########################################################################
#########################################################################
#########################################################################

#########################################################################
#########################################################################
# REQUIRED READING:
#
# Practical advice / notes / basic principles on statistical model 
#    comparison in general, and in BioGeoBEARS:
# http://phylo.wikidot.com/advice-on-statistical-model-comparison-in-biogeobears
#########################################################################
#########################################################################

# Set up empty tables to hold the statistical results
restable = NULL
teststable = NULL

#######################################################
# Statistics -- DEC vs. DEC+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDEC)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDECj)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DEC, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDEC, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DEC+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECj, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

# The null hypothesis for a Likelihood Ratio Test (LRT) is that two models
# confer the same likelihood on the data. See: Brian O'Meara's webpage:
# http://www.brianomeara.info/tutorials/aic
# ...for an intro to LRT, AIC, and AICc

rbind(res2, res1)
tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#######################################################
# Statistics -- DIVALIKE vs. DIVALIKE+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKE)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEj)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DIVALIKE, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKE, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DIVALIKE+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEj, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

rbind(res2, res1)
conditional_format_table(stats)

tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)


#########################################################################
# RESULTS: DEC, DEC+J, DIVALIKE, DIVALIKE+J
#########################################################################
teststable$alt = c("DEC+J", "DIVALIKE+J")
teststable$null = c("DEC", "DIVALIKE")
row.names(restable) = c("DEC", "DEC+J", "DIVALIKE", "DIVALIKE+J")

# Look at the results!!
restable
teststable

#######################################################
# Save the results tables for later -- check for e.g.
# convergence issues
#######################################################

# Loads to "restable"
save(restable, file="restable_v1.Rdata")
load(file="restable_v1.Rdata")

# Loads to "teststable"
save(teststable, file="teststable_v1.Rdata")
load(file="teststable_v1.Rdata")

# Also save to text files
write.table(restable, file="restable.txt", quote=FALSE, sep="\t")
write.table(unlist_df(teststable), file="teststable.txt", quote=FALSE, sep="\t")

#######################################################
# Model weights of all six models
#######################################################
restable2 = restable

# With AICs:
AICtable = calc_AIC_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams)
restable = cbind(restable, AICtable)
restable_AIC_rellike = AkaikeWeights_on_summary_table(restable=restable, colname_to_use="AIC")
restable_AIC_rellike

# With AICcs -- factors in sample size
samplesize = length(tr$tip.label)
AICtable = calc_AICc_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams, samplesize=samplesize)
restable2 = cbind(restable2, AICtable)
restable_AICc_rellike = AkaikeWeights_on_summary_table(restable=restable2, colname_to_use="AIC")
restable_AICc_rellike
free_params = row.names(resDECj$output@params_table[resDECj$output@params_table$type=="free",])
names(restable_AICc_rellike) = c("LnL", "numparams", free_params, "AICc", "AICc_wt")

# Also save to text files
write.table(restable_AIC_rellike, file="restable_AIC_rellike.txt", quote=FALSE, sep="\t")
write.table(restable_AICc_rellike, file="restable_AICc_rellike.txt", quote=FALSE, sep="\t")

# Save with nice conditional formatting
write.table(conditional_format_table(restable_AIC_rellike), file="restable_AIC_rellike_formatted.txt", quote=FALSE, sep="\t")
write.table(conditional_format_table(restable_AICc_rellike), file="restable_AICc_rellike_formatted.txt", quote=FALSE, sep="\t")


```




10. Now we are going to use the model DEC with dispersal multipliers

```{r}

#######################################################
# Run DEC
#######################################################

# Intitialize a default model (DEC model)
BioGeoBEARS_run_object = define_BioGeoBEARS_run()

# Give BioGeoBEARS the location of the phylogeny Newick file
BioGeoBEARS_run_object$trfn = trfn

# Give BioGeoBEARS the location of the geography text file
BioGeoBEARS_run_object$geogfn = geogfn

# Input the maximum range size
BioGeoBEARS_run_object$max_range_size = max_range_size

BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
# 1. Here, un-comment ONLY the files you want to use.
# 2. Also un-comment "BioGeoBEARS_run_object = section_the_tree(...", below.
# 3. For example files see (a) extdata_dir, 
#  or (b) http://phylo.wikidot.com/biogeobears#files
#  and BioGeoBEARS Google Group posts for further hints)
#
# Uncomment files you wish to use in time-stratified analyses:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = areasadjacency
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
# (use more cores to speed it up; this requires
# library(parallel) and/or library(snow). The package "parallel" 
# is now default on Macs in R 3.0+, but apparently still 
# has to be typed on some Windows machines. Note: apparently 
# parallel works on Mac command-line R, but not R.app.
# BioGeoBEARS checks for this and resets to 1
# core with R.app)

# Sparse matrix exponentiation is an option for huge numbers of ranges/states (600+)
# I have experimented with sparse matrix exponentiation in EXPOKIT/rexpokit,
# but the results are imprecise and so I haven't explored it further.
# In a Bayesian analysis, it might work OK, but the ML point estimates are
# not identical.
# Also, I have not implemented all functions to work with force_sparse=TRUE.
# Volunteers are welcome to work on it!!
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC model
# (nothing to do; defaults)

# Look at the BioGeoBEARS_run_object; it's just a list of settings etc.
BioGeoBEARS_run_object

# This contains the model object
BioGeoBEARS_run_object$BioGeoBEARS_model_object

# This table contains the parameters of the model 
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table

# Run this to check inputs. Read the error messages if you get them!
check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "Glareosae_DEC_M0_constrained_v1.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDECc = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECc = res
    }



```

11. Run DEC + J con tasas de dispersion

```{r}



#######################################################
# Run DEC+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDEC$outputs@params_table["d","est"]
estart = resDEC$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Add j as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
resfn = "Glareosae_DEC+J_M0_constrained_v1.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDECjc = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECjc = res
    }
```


12. Plotting the results with dispesal rates

```{r}

#######################################################
# PDF plots
#######################################################
pdffn = "Glareosae_DEC_vs_DEC+J_M0_constrained_v1.pdf"
pdf(pdffn, width=6, height=6)

#######################################################
# Plot ancestral states - DEC
#######################################################
analysis_titletxt ="BioGeoBEARS DEC on Glareosae M0_constrained"

# Setup
results_object = resDECc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DECJ
#######################################################
analysis_titletxt ="BioGeoBEARS DEC+J on Glareosae M0_constrained"

# Setup
results_object = resDECjc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()  # Turn off PDF
cmdstr = paste("open ", pdffn, sep="")
system(cmdstr) # Plot it

```

13. Run DIVA like with dispersal rates

```{r}
#######################################################
# Run DIVALIKE
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE model
# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# No jump dispersal/founder-event speciation
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = 0.01
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = 0.01

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

runslow = TRUE
resfn = "Glareosae_DIVALIKE_M0_constrained_v1.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDIVALIKEc = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEc = res
    }

```

14. Run DIVALIKE + j with dispersal rates

```{r}
#######################################################
# Run DIVALIKE+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDIVALIKE$outputs@params_table["d","est"]
estart = resDIVALIKE$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add jump dispersal/founder-event speciation
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Under DIVALIKE+J, the max of "j" should be 2, not 3 (as is default in DEC+J)
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","min"] = 0.00001
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","max"] = 1.99999

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

resfn = "Glareosae_DIVALIKE+J_M0_constrained_v1.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDIVALIKEjc = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEjc = res
    }


```

15. Plotting additional results with dispersal rates

```{r}

pdffn = "Glareosae_DIVALIKE_vs_DIVALIKE+J_M0_constrained_v1.pdf"
pdf(pdffn, width=6, height=6)

#######################################################
# Plot ancestral states - DIVALIKE
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE on Glareosae M0_constrained"

# Setup
results_object = resDIVALIKEc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DIVALIKE+J
#######################################################
analysis_titletxt ="Glareosae DIVALIKE+J on Psychotria M0_constrained"

# Setup
results_object = resDIVALIKEjc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()
cmdstr = paste("open ", pdffn, sep="")
system(cmdstr)

```


16. Comparing all results with dispersal rates

```{r}
#########################################################################
#########################################################################
#########################################################################
#########################################################################
# 
# CALCULATE SUMMARY STATISTICS TO COMPARE
# DEC, DEC+J, DIVALIKE, DIVALIKE+J
# 
#########################################################################
#########################################################################
#########################################################################
#########################################################################

#########################################################################
#########################################################################
# REQUIRED READING:
#
# Practical advice / notes / basic principles on statistical model 
#    comparison in general, and in BioGeoBEARS:
# http://phylo.wikidot.com/advice-on-statistical-model-comparison-in-biogeobears
#########################################################################
#########################################################################

# Set up empty tables to hold the statistical results
restable = NULL
teststable = NULL

#######################################################
# Statistics -- DEC vs. DEC+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDECc)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDECjc)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DEC, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECc, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DEC+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECjc, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

# The null hypothesis for a Likelihood Ratio Test (LRT) is that two models
# confer the same likelihood on the data. See: Brian O'Meara's webpage:
# http://www.brianomeara.info/tutorials/aic
# ...for an intro to LRT, AIC, and AICc

rbind(res2, res1)
tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#######################################################
# Statistics -- DIVALIKE vs. DIVALIKE+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEc)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEjc)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DIVALIKE, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEc, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DIVALIKE+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEjc, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

rbind(res2, res1)
conditional_format_table(stats)

tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#########################################################################
# RESULTS: DEC, DEC+J, DIVALIKE, DIVALIKE+J
#########################################################################
teststable$alt = c("DEC+Jc", "DIVALIKE+Jc")
teststable$null = c("DECc", "DIVALIKEc")
row.names(restable) = c("DECc", "DEC+Jc", "DIVALIKEc", "DIVALIKE+Jc")

# Look at the results!!
restable
teststable

#######################################################
# Save the results tables for later -- check for e.g.
# convergence issues
#######################################################

# Loads to "restable"
save(restable, file="restable_v1c.Rdata")
load(file="restable_v1c.Rdata")

# Loads to "teststable"
save(teststable, file="teststable_v1c.Rdata")
load(file="teststable_v1c.Rdata")

# Also save to text files
write.table(restable, file="restablec.txt", quote=FALSE, sep="\t")
write.table(unlist_df(teststable), file="teststablec.txt", quote=FALSE, sep="\t")

#######################################################
# Model weights of all six models
#######################################################
restable2 = restable

# With AICs:
AICtable = calc_AIC_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams)
restable = cbind(restable, AICtable)
restable_AIC_rellike = AkaikeWeights_on_summary_table(restable=restable, colname_to_use="AIC")
restable_AIC_rellike

# With AICcs -- factors in sample size
samplesize = length(tr$tip.label)
AICtable = calc_AICc_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams, samplesize=samplesize)
restable2 = cbind(restable2, AICtable)
restable_AICc_rellike = AkaikeWeights_on_summary_table(restable=restable2, colname_to_use="AIC")
restable_AICc_rellike
free_params = row.names(resDECj$output@params_table[resDECj$output@params_table$type=="free",])
names(restable_AICc_rellike) = c("LnL", "numparams", free_params, "AICc", "AICc_wt")

# Also save to text files
write.table(restable_AIC_rellike, file="restable_AIC_rellikec.txt", quote=FALSE, sep="\t")
write.table(restable_AICc_rellike, file="restable_AICc_rellikec.txt", quote=FALSE, sep="\t")

# Save with nice conditional formatting
write.table(conditional_format_table(restable_AIC_rellike), file="restable_AIC_rellike_formattedc.txt", quote=FALSE, sep="\t")
write.table(conditional_format_table(restable_AICc_rellike), file="restable_AICc_rellike_formattedc.txt", quote=FALSE, sep="\t")


```


17. Now we are going to use the model DEC with dispersal multipliers bis

```{r}

#######################################################
# Run DEC
#######################################################

# Intitialize a default model (DEC model)
BioGeoBEARS_run_object = define_BioGeoBEARS_run()

# Give BioGeoBEARS the location of the phylogeny Newick file
BioGeoBEARS_run_object$trfn = trfn

# Give BioGeoBEARS the location of the geography text file
BioGeoBEARS_run_object$geogfn = geogfn

# Input the maximum range size
BioGeoBEARS_run_object$max_range_size = max_range_size

BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
# 1. Here, un-comment ONLY the files you want to use.
# 2. Also un-comment "BioGeoBEARS_run_object = section_the_tree(...", below.
# 3. For example files see (a) extdata_dir, 
#  or (b) http://phylo.wikidot.com/biogeobears#files
#  and BioGeoBEARS Google Group posts for further hints)
#
# Uncomment files you wish to use in time-stratified analyses:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers_b.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = areasadjacency
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
# (use more cores to speed it up; this requires
# library(parallel) and/or library(snow). The package "parallel" 
# is now default on Macs in R 3.0+, but apparently still 
# has to be typed on some Windows machines. Note: apparently 
# parallel works on Mac command-line R, but not R.app.
# BioGeoBEARS checks for this and resets to 1
# core with R.app)

# Sparse matrix exponentiation is an option for huge numbers of ranges/states (600+)
# I have experimented with sparse matrix exponentiation in EXPOKIT/rexpokit,
# but the results are imprecise and so I haven't explored it further.
# In a Bayesian analysis, it might work OK, but the ML point estimates are
# not identical.
# Also, I have not implemented all functions to work with force_sparse=TRUE.
# Volunteers are welcome to work on it!!
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC model
# (nothing to do; defaults)

# Look at the BioGeoBEARS_run_object; it's just a list of settings etc.
BioGeoBEARS_run_object

# This contains the model object
BioGeoBEARS_run_object$BioGeoBEARS_model_object

# This table contains the parameters of the model 
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table

# Run this to check inputs. Read the error messages if you get them!
check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "Glareosae_DEC_M0_constrained_b_v1.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDECcb = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECcb = res
    }



```

18. Run DEC + J con tasas de dispersion bis

```{r}



#######################################################
# Run DEC+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers_b.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDEC$outputs@params_table["d","est"]
estart = resDEC$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Add j as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
resfn = "Glareosae_DEC+J_M0_constrained_b_v1.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDECjcb = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECjcb = res
    }
```


19. Plotting the results with dispesal rates

```{r}

#######################################################
# PDF plots
#######################################################
pdffn = "Glareosae_DEC_vs_DEC+J_M0_constrained_b_v1.pdf"
pdf(pdffn, width=6, height=6)

#######################################################
# Plot ancestral states - DEC
#######################################################
analysis_titletxt ="BioGeoBEARS DEC on Glareosae M0_constrained_b"

# Setup
results_object = resDECcb
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DECJ
#######################################################
analysis_titletxt ="BioGeoBEARS DEC+J on Glareosae M0_constrained_b"

# Setup
results_object = resDECjcb
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()  # Turn off PDF
cmdstr = paste("open ", pdffn, sep="")
system(cmdstr) # Plot it

```

20. Run DIVA like with dispersal rates bis

```{r}
#######################################################
# Run DIVALIKE
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers_b.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE model
# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# No jump dispersal/founder-event speciation
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = 0.01
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = 0.01

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

runslow = TRUE
resfn = "Glareosae_DIVALIKE_M0_constrained_b_v1.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDIVALIKEcb = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEcb = res
    }

```

21. Run DIVALIKE + j with dispersal rates bis

```{r}
#######################################################
# Run DIVALIKE+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers_b.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDIVALIKE$outputs@params_table["d","est"]
estart = resDIVALIKE$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add jump dispersal/founder-event speciation
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Under DIVALIKE+J, the max of "j" should be 2, not 3 (as is default in DEC+J)
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","min"] = 0.00001
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","max"] = 1.99999

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

resfn = "Glareosae_DIVALIKE+J_M0_constrained_b_v1.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDIVALIKEjcb = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEjcb = res
    }


```

22. Plotting additional results with dispersal rates bis

```{r}

pdffn = "Glareosae_DIVALIKE_vs_DIVALIKE+J_M0_constrained_b_v1.pdf"
pdf(pdffn, width=6, height=6)

#######################################################
# Plot ancestral states - DIVALIKE
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE on Glareosae M0_constrained_b"

# Setup
results_object = resDIVALIKEcb
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DIVALIKE+J
#######################################################
analysis_titletxt ="Glareosae DIVALIKE+J on Psychotria M0_constrained_b"

# Setup
results_object = resDIVALIKEjcb
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()
cmdstr = paste("open ", pdffn, sep="")
system(cmdstr)

```


23. Comparing all results with dispersal rates bis

```{r}
#########################################################################
#########################################################################
#########################################################################
#########################################################################
# 
# CALCULATE SUMMARY STATISTICS TO COMPARE
# DEC, DEC+J, DIVALIKE,DIVALIKE+J
# 
#########################################################################
#########################################################################
#########################################################################
#########################################################################

#########################################################################
#########################################################################
# REQUIRED READING:
#
# Practical advice / notes / basic principles on statistical model 
#    comparison in general, and in BioGeoBEARS:
# http://phylo.wikidot.com/advice-on-statistical-model-comparison-in-biogeobears
#########################################################################
#########################################################################

# Set up empty tables to hold the statistical results
restable = NULL
teststable = NULL

#######################################################
# Statistics -- DEC vs. DEC+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDECcb)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDECjcb)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DEC, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECcb, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DEC+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECjcb, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

# The null hypothesis for a Likelihood Ratio Test (LRT) is that two models
# confer the same likelihood on the data. See: Brian O'Meara's webpage:
# http://www.brianomeara.info/tutorials/aic
# ...for an intro to LRT, AIC, and AICc

rbind(res2, res1)
tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#######################################################
# Statistics -- DIVALIKE vs. DIVALIKE+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEcb)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEjcb)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DIVALIKE, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEcb, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DIVALIKE+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEjcb, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

rbind(res2, res1)
conditional_format_table(stats)

tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#########################################################################
# RESULTS: DEC, DEC+J, DIVALIKE, DIVALIKE+J
#########################################################################
teststable$alt = c("DEC+Jc", "DIVALIKE+Jc")
teststable$null = c("DECc", "DIVALIKEc")
row.names(restable) = c("DECc", "DEC+Jc", "DIVALIKEc", "DIVALIKE+Jc")

# Look at the results!!
restable
teststable

#######################################################
# Save the results tables for later -- check for e.g.
# convergence issues
#######################################################

# Loads to "restable"
save(restable, file="restable_v1cb.Rdata")
load(file="restable_v1cb.Rdata")

# Loads to "teststable"
save(teststable, file="teststable_v1cb.Rdata")
load(file="teststable_v1cb.Rdata")

# Also save to text files
write.table(restable, file="restablecb.txt", quote=FALSE, sep="\t")
write.table(unlist_df(teststable), file="teststablecb.txt", quote=FALSE, sep="\t")

#######################################################
# Model weights of all six models
#######################################################
restable2 = restable

# With AICs:
AICtable = calc_AIC_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams)
restable = cbind(restable, AICtable)
restable_AIC_rellike = AkaikeWeights_on_summary_table(restable=restable, colname_to_use="AIC")
restable_AIC_rellike

# With AICcs -- factors in sample size
samplesize = length(tr$tip.label)
AICtable = calc_AICc_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams, samplesize=samplesize)
restable2 = cbind(restable2, AICtable)
restable_AICc_rellike = AkaikeWeights_on_summary_table(restable=restable2, colname_to_use="AIC")
restable_AICc_rellike
free_params = row.names(resDECjcb$output@params_table[resDECjcb$output@params_table$type=="free",])
names(restable_AICc_rellike) = c("LnL", "numparams", free_params, "AICc", "AICc_wt")

# Also save to text files
write.table(restable_AIC_rellike, file="restable_AIC_rellikecb.txt", quote=FALSE, sep="\t")
write.table(restable_AICc_rellike, file="restable_AICc_rellikecb.txt", quote=FALSE, sep="\t")

# Save with nice conditional formatting
write.table(conditional_format_table(restable_AIC_rellike), file="restable_AIC_rellike_formattedcb.txt", quote=FALSE, sep="\t")
write.table(conditional_format_table(restable_AICc_rellike), file="restable_AICc_rellike_formattedcb.txt", quote=FALSE, sep="\t")


```


24. Now we are going to use the model DEC with dispersal multipliers and time periods

```{r}

#######################################################
# Run DEC
#######################################################

# Intitialize a default model (DEC model)
BioGeoBEARS_run_object = define_BioGeoBEARS_run()

# Give BioGeoBEARS the location of the phylogeny Newick file
BioGeoBEARS_run_object$trfn = trfn

# Give BioGeoBEARS the location of the geography text file
BioGeoBEARS_run_object$geogfn = geogfn

# Input the maximum range size
BioGeoBEARS_run_object$max_range_size = max_range_size

BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
# 1. Here, un-comment ONLY the files you want to use.
# 2. Also un-comment "BioGeoBEARS_run_object = section_the_tree(...", below.
# 3. For example files see (a) extdata_dir, 
#  or (b) http://phylo.wikidot.com/biogeobears#files
#  and BioGeoBEARS Google Group posts for further hints)
#
# Uncomment files you wish to use in time-stratified analyses:
BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers2.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = areasadjacency
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
# (use more cores to speed it up; this requires
# library(parallel) and/or library(snow). The package "parallel" 
# is now default on Macs in R 3.0+, but apparently still 
# has to be typed on some Windows machines. Note: apparently 
# parallel works on Mac command-line R, but not R.app.
# BioGeoBEARS checks for this and resets to 1
# core with R.app)

# Sparse matrix exponentiation is an option for huge numbers of ranges/states (600+)
# I have experimented with sparse matrix exponentiation in EXPOKIT/rexpokit,
# but the results are imprecise and so I haven't explored it further.
# In a Bayesian analysis, it might work OK, but the ML point estimates are
# not identical.
# Also, I have not implemented all functions to work with force_sparse=TRUE.
# Volunteers are welcome to work on it!!
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC model
# (nothing to do; defaults)

# Look at the BioGeoBEARS_run_object; it's just a list of settings etc.
BioGeoBEARS_run_object

# This contains the model object
BioGeoBEARS_run_object$BioGeoBEARS_model_object

# This table contains the parameters of the model 
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table

# Run this to check inputs. Read the error messages if you get them!
check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "Glareosae_DEC_M0_constrained2_v1.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDECc2 = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECc2 = res
    }



```

25. Run DEC + J with distance multipliers and time periods

```{r}



#######################################################
# Run DEC+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers2.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDEC$outputs@params_table["d","est"]
estart = resDEC$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Add j as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
resfn = "Glareosae_DEC+J_M0_constrained2_v1.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDECjc2 = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECjc2 = res
    }
```


26. Plotting the results with dispesal rates and time periods

```{r}

#######################################################
# PDF plots
#######################################################
pdffn = "Glareosae_DEC_vs_DEC+J_M0_constrained2_v1.pdf"
pdf(pdffn, width=6, height=6)

#######################################################
# Plot ancestral states - DEC
#######################################################
analysis_titletxt ="BioGeoBEARS DEC on Glareosae M0_constrained and time periods"

# Setup
results_object = resDECc2
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DECJ
#######################################################
analysis_titletxt ="BioGeoBEARS DEC+J on Glareosae M0_constrained and time periods"

# Setup
results_object = resDECjc2
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()  # Turn off PDF
cmdstr = paste("open ", pdffn, sep="")
system(cmdstr) # Plot it

```

27. Run DIVA like with dispersal rates and time periods

```{r}
#######################################################
# Run DIVALIKE
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers2.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE model
# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# No jump dispersal/founder-event speciation
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = 0.01
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = 0.01

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

runslow = TRUE
resfn = "Glareosae_DIVALIKE_M0_constrained2_v1.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDIVALIKEc2 = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEc2 = res
    }

```

28. Run DIVALIKE + j with dispersal rates and time periods

```{r}
#######################################################
# Run DIVALIKE+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers2.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDIVALIKE$outputs@params_table["d","est"]
estart = resDIVALIKE$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add jump dispersal/founder-event speciation
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Under DIVALIKE+J, the max of "j" should be 2, not 3 (as is default in DEC+J)
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","min"] = 0.00001
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","max"] = 1.99999

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

resfn = "Glareosae_DIVALIKE+J_M0_constrained2_v1.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDIVALIKEjc2 = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEjc2 = res
    }


```

29. Plotting additional results with dispersal rates and time periods

```{r}

pdffn = "Glareosae_DIVALIKE_vs_DIVALIKE+J_M0_constrained2_v1.pdf"
pdf(pdffn, width=6, height=6)

#######################################################
# Plot ancestral states - DIVALIKE
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE on Glareosae M0_constrained and time periods"

# Setup
results_object = resDIVALIKEc2
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DIVALIKE+J
#######################################################
analysis_titletxt ="Glareosae DIVALIKE+J on Psychotria M0_constrained and time periods"

# Setup
results_object = resDIVALIKEjc2
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()
cmdstr = paste("open ", pdffn, sep="")
system(cmdstr)

```


30. Comparing all results with dispersal rates and time periods

```{r}
#########################################################################
#########################################################################
#########################################################################
#########################################################################
# 
# CALCULATE SUMMARY STATISTICS TO COMPARE
# DEC, DEC+J, DIVALIKE, DIVALIKE+J
# 
#########################################################################
#########################################################################
#########################################################################
#########################################################################

#########################################################################
#########################################################################
# REQUIRED READING:
#
# Practical advice / notes / basic principles on statistical model 
#    comparison in general, and in BioGeoBEARS:
# http://phylo.wikidot.com/advice-on-statistical-model-comparison-in-biogeobears
#########################################################################
#########################################################################

# Set up empty tables to hold the statistical results
restable = NULL
teststable = NULL

#######################################################
# Statistics -- DEC vs. DEC+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDECc2)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDECjc2)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DEC, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECc2, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DEC+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECjc2, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

# The null hypothesis for a Likelihood Ratio Test (LRT) is that two models
# confer the same likelihood on the data. See: Brian O'Meara's webpage:
# http://www.brianomeara.info/tutorials/aic
# ...for an intro to LRT, AIC, and AICc

rbind(res2, res1)
tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#######################################################
# Statistics -- DIVALIKE vs. DIVALIKE+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEc2)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEjc2)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DIVALIKE, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEc2, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DIVALIKE+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEjc2, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

rbind(res2, res1)
conditional_format_table(stats)

tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)


#########################################################################
# RESULTS: DEC, DEC+J, DIVALIKE, DIVALIKE+J
#########################################################################
teststable$alt = c("DEC+Jc", "DIVALIKE+Jc")
teststable$null = c("DECc", "DIVALIKEc")
row.names(restable) = c("DECc", "DEC+Jc", "DIVALIKEc", "DIVALIKE+Jc")

# Look at the results!!
restable
teststable

#######################################################
# Save the results tables for later -- check for e.g.
# convergence issues
#######################################################

# Loads to "restable"
save(restable, file="restable_v1c2.Rdata")
load(file="restable_v1c2.Rdata")

# Loads to "teststable"
save(teststable, file="teststable_v1c2.Rdata")
load(file="teststable_v1c2.Rdata")

# Also save to text files
write.table(restable, file="restablec2.txt", quote=FALSE, sep="\t")
write.table(unlist_df(teststable), file="teststablec2.txt", quote=FALSE, sep="\t")

#######################################################
# Model weights of all six models
#######################################################
restable2 = restable

# With AICs:
AICtable = calc_AIC_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams)
restable = cbind(restable, AICtable)
restable_AIC_rellike = AkaikeWeights_on_summary_table(restable=restable, colname_to_use="AIC")
restable_AIC_rellike

# With AICcs -- factors in sample size
samplesize = length(tr$tip.label)
AICtable = calc_AICc_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams, samplesize=samplesize)
restable2 = cbind(restable2, AICtable)
restable_AICc_rellike = AkaikeWeights_on_summary_table(restable=restable2, colname_to_use="AIC")
restable_AICc_rellike
free_params = row.names(resDECj$output@params_table[resDECj$output@params_table$type=="free",])
names(restable_AICc_rellike) = c("LnL", "numparams", free_params, "AICc", "AICc_wt")

# Also save to text files
write.table(restable_AIC_rellike, file="restable_AIC_rellikec2.txt", quote=FALSE, sep="\t")
write.table(restable_AICc_rellike, file="restable_AICc_rellikec2.txt", quote=FALSE, sep="\t")

# Save with nice conditional formatting
write.table(conditional_format_table(restable_AIC_rellike), file="restable_AIC_rellike_formattedc2.txt", quote=FALSE, sep="\t")
write.table(conditional_format_table(restable_AICc_rellike), file="restable_AICc_rellike_formattedc2.txt", quote=FALSE, sep="\t")


```



31. Setting packages and sources for multiples trees analyses

```{r}
###########################################################
# Run Biogeobears on multiple trees and store the output 
###########################################################

###########################################################
#Load libraries and additional source code
###########################################################
library(optimx)         
library(FD)       
library(snow)     
library(parallel)
library(BioGeoBEARS)
library(stringr)

source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_add_fossils_randomly_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_basics_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_calc_transition_matrices_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_classes_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_detection_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_DNA_cladogenesis_sim_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_extract_Qmat_COOmat_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_generics_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_models_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_on_multiple_trees_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_plots_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_readwrite_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_simulate_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_SSEsim_makePlots_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_SSEsim_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_stochastic_mapping_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_stratified_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_univ_model_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/calc_uppass_probs_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/calc_loglike_sp_v01.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/get_stratified_subbranch_top_downpass_likelihoods_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/runBSM_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/stochastic_map_given_inputs.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/summarize_BSM_tables_v1.R")

calc_loglike_sp = compiler::cmpfun(calc_loglike_sp_prebyte)    # crucial to fix bug in uppass calculations
calc_independent_likelihoods_on_each_branch = compiler::cmpfun(calc_independent_likelihoods_on_each_branch_prebyte)



```

32. Setting files for multiple trees analyses

```{r}


newick_fns = read.nexus("tree.trees")
newick_fns
tipstoremove <- c("Cremota","Cspicata", "Cpaniculata", "Celongata", "Cscoparia", "Cmerritt-fernaldi", "Cbolanderi", "Cleptopoda", "Cdeweyana", "Cbromoides", "Cchordorrhyza")
newick_fns2 <- list()
for (i in 1:50) newick_fns2[[i]] <- drop.tip (newick_fns[[i]], tipstoremove)
class(newick_fns2) <- "multiPhylo"

plot(newick_fns2[[4]], cex = 0.8)
title("Example Glareosae phylogeny")
axisPhylo()

plot(newick_fns2[[4]], cex = 0.8, root.edge = TRUE)
title("Example Glareosae phylogeny")
axisPhylo()

write.tree(newick_fns2, file = "glareosae.trees")
write.tree(newick_fns2[[1]], file = "glareosae1.newick")
write.tree(newick_fns2[[2]], file = "glareosae2.newick")
write.tree(newick_fns2[[3]], file = "glareosae3.newick")
write.tree(newick_fns2[[4]], file = "glareosae4.newick")
write.tree(newick_fns2[[5]], file = "glareosae5.newick")
write.tree(newick_fns2[[6]], file = "glareosae6.newick")
write.tree(newick_fns2[[7]], file = "glareosae7.newick")
write.tree(newick_fns2[[8]], file = "glareosae8.newick")
write.tree(newick_fns2[[9]], file = "glareosae9.newick")
write.tree(newick_fns2[[10]], file = "glareosae10.newick")
write.tree(newick_fns2[[11]], file = "glareosae11.newick")
write.tree(newick_fns2[[12]], file = "glareosae12.newick")
write.tree(newick_fns2[[13]], file = "glareosae13.newick")
write.tree(newick_fns2[[14]], file = "glareosae14.newick")
write.tree(newick_fns2[[15]], file = "glareosae15.newick")
write.tree(newick_fns2[[16]], file = "glareosae16.newick")
write.tree(newick_fns2[[17]], file = "glareosae17.newick")
write.tree(newick_fns2[[18]], file = "glareosae18.newick")
write.tree(newick_fns2[[19]], file = "glareosae19.newick")
write.tree(newick_fns2[[20]], file = "glareosae20.newick")
write.tree(newick_fns2[[21]], file = "glareosae21.newick")
write.tree(newick_fns2[[22]], file = "glareosae22.newick")
write.tree(newick_fns2[[23]], file = "glareosae23.newick")
write.tree(newick_fns2[[24]], file = "glareosae24.newick")
write.tree(newick_fns2[[25]], file = "glareosae25.newick")
write.tree(newick_fns2[[26]], file = "glareosae26.newick")
write.tree(newick_fns2[[27]], file = "glareosae27.newick")
write.tree(newick_fns2[[28]], file = "glareosae28.newick")
write.tree(newick_fns2[[29]], file = "glareosae29.newick")
write.tree(newick_fns2[[30]], file = "glareosae30.newick")
write.tree(newick_fns2[[31]], file = "glareosae31.newick")
write.tree(newick_fns2[[32]], file = "glareosae32.newick")
write.tree(newick_fns2[[33]], file = "glareosae33.newick")
write.tree(newick_fns2[[34]], file = "glareosae34.newick")
write.tree(newick_fns2[[35]], file = "glareosae35.newick")
write.tree(newick_fns2[[36]], file = "glareosae36.newick")
write.tree(newick_fns2[[37]], file = "glareosae37.newick")
write.tree(newick_fns2[[38]], file = "glareosae38.newick")
write.tree(newick_fns2[[39]], file = "glareosae39.newick")
write.tree(newick_fns2[[40]], file = "glareosae40.newick")
write.tree(newick_fns2[[41]], file = "glareosae41.newick")
write.tree(newick_fns2[[42]], file = "glareosae42.newick")
write.tree(newick_fns2[[43]], file = "glareosae43.newick")
write.tree(newick_fns2[[44]], file = "glareosae44.newick")
write.tree(newick_fns2[[45]], file = "glareosae45.newick")
write.tree(newick_fns2[[46]], file = "glareosae46.newick")
write.tree(newick_fns2[[47]], file = "glareosae47.newick")
write.tree(newick_fns2[[48]], file = "glareosae48.newick")
write.tree(newick_fns2[[49]], file = "glareosae49.newick")
write.tree(newick_fns2[[50]], file = "glareosae50.newick")


newick_fns = scan("glareosae.newicks", what="character", sep=NULL)
geog_fns = scan("glareosae.datas", what="character", sep=NULL)
resfns = scan("outputnames.txt", what="character", sep=NULL)





trfn = np("glareosae.tree")
moref(trfn)

geogfn = np("glareosae.data")
#extdata_dir = np(system.file("extdata", package="BioGeoBEARS"))
#extdata_dir
#list.files(extdata_dir)
#geogfn = np(paste(addslash(extdata_dir), "Psychotria_geog.data", sep=""))

geogfn
moref(geogfn)

# Look at your geographic range data:
tipranges = getranges_from_LagrangePHYLIP(lgdata_fn=geogfn)
tipranges

max_range_size = 5




numstates_from_numareas(numareas=5, maxareas=5, include_null_range=TRUE)
numstates_from_numareas(numareas=5, maxareas=5, include_null_range=FALSE)



```

33. Running DEC on multiples trees


```{r}


#######################################################
# Run DEC
#######################################################

# Intitialize a default model (DEC model)
BioGeoBEARS_run_object = define_BioGeoBEARS_run()

# Give BioGeoBEARS the location of the phylogeny Newick file
BioGeoBEARS_run_object$trfn = trfn

# Give BioGeoBEARS the location of the geography text file
BioGeoBEARS_run_object$geogfn = geogfn

# Input the maximum range size
BioGeoBEARS_run_object$max_range_size = max_range_size

BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
# 1. Here, un-comment ONLY the files you want to use.
# 2. Also un-comment "BioGeoBEARS_run_object = section_the_tree(...", below.
# 3. For example files see (a) extdata_dir, 
#  or (b) http://phylo.wikidot.com/biogeobears#files
#  and BioGeoBEARS Google Group posts for further hints)
#
# Uncomment files you wish to use in time-stratified analyses:
BioGeoBEARS_run_object$timesfn = "timeperiods2.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers2.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = areasadjacency
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
# (use more cores to speed it up; this requires
# library(parallel) and/or library(snow). The package "parallel" 
# is now default on Macs in R 3.0+, but apparently still 
# has to be typed on some Windows machines. Note: apparently 
# parallel works on Mac command-line R, but not R.app.
# BioGeoBEARS checks for this and resets to 1
# core with R.app)

# Sparse matrix exponentiation is an option for huge numbers of ranges/states (600+)
# I have experimented with sparse matrix exponentiation in EXPOKIT/rexpokit,
# but the results are imprecise and so I haven't explored it further.
# In a Bayesian analysis, it might work OK, but the ML point estimates are
# not identical.
# Also, I have not implemented all functions to work with force_sparse=TRUE.
# Volunteers are welcome to work on it!!
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC model
# (nothing to do; defaults)

# Look at the BioGeoBEARS_run_object; it's just a list of settings etc.
BioGeoBEARS_run_object

# This contains the model object
BioGeoBEARS_run_object$BioGeoBEARS_model_object

# This table contains the parameters of the model 
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table

# Run this to check inputs. Read the error messages if you get them!
check_BioGeoBEARS_run(BioGeoBEARS_run_object)




############################################################
##
#NOTE: WE ARE USING POINT 34 WHICH HAS BEEN UPDATED IN THE GOOGLE GROUP. DO NOT USE THE NUMBER !!!!!!!!!!!!!!!!!!!


res = run_bears_optim_on_multiple_trees(BioGeoBEARS_run_object, newick_fns, model_name="Glareosae_DECc2_MT", geog_fns, resfns, run_or_collect="run", start_treenum=1, end_treenum=length(newick_fns), runslow=TRUE, plot_params=FALSE)
res

res2 = run_bears_optim_on_multiple_trees(BioGeoBEARS_run_object, newick_fns, model_name="Glareosae_DECc2_MT", geog_fns, resfns, run_or_collect="collect", start_treenum=1, end_treenum=length(newick_fns), runslow=TRUE, plot_params=FALSE)
res2


names(res2)

# This has the (averaged) state probabilities at nodes and corners
# for the master tree (whatever you put in BioGeoBEARS_run_object)
dim(res2$state_probs_at_nodes_across_all_trees)
dim(res2$state_probs_at_corners_across_all_trees)

# This has the averaged parameter values
res2$optim_results_mean





# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "Glareosae_DEC_M0_constrained2_v1_MT.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDECc2MT = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECc2MT = res
    }



```


34. ANALYSES WITH MULTIPLE TREES DOWNLOADED FROM THE GOOGLE GROUP
WE HAVE SET THE ANALYSES USING COMMANDS FROM NUMBER 33 BUT THE FINAL RUN IS DONE HERE.

```{r}

###########################################################
# Run BioGeoBEARS on multiple trees and store the output 
#- mostrar texto citado -
tipranges = getranges_from_LagrangePHYLIP(lgdata_fn=geogfn)
tipranges
#- mostrar texto citado -
# Run the analysis, on the master tree
# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "master_tree_single_run_v1.Rdata"
if (runslow)
    {
    master_tree_res = bears_optim_run(BioGeoBEARS_run_object)
    master_tree_res

    save(master_tree_res, file=resfn)
    resDEC = master_tree_res
    } else {
    # Loads to "master_tree_res"
    load(resfn)
    resDEC = master_tree_res
    }


# multiple tree settings
#' @BioGeoBEARS_run_object Set up the inference model you want to run on 
#' each tree, like you would for a normal single-tree run.
#' @newick_fns A list of the Newick files (e.g., you should extract some trees
#' from a BEAST NEXUS MCMC output)
#' @model_name The name you would like added to output filenames
#' @geog_fns A list of corresponding geography files (by default, these are just .geog instead of .newick)
#' @resfns A list of results filenames for .Rdata files, either for saving BioGeoBEARS analyses on each tree, or 
#' for loading previously-saved runs (by default, these are just _model_name.Rdata instead of .newick)
#' @run_or_collect If you want to run BioGeoBEARS on each tree (slow), use "run". If you just want to 
#' collect the results over all the trees, use "collect".  For both, pick "both".
#' @start_treenum Default 1. Change if you want to skip some trees
#' @end_treenum Default is length(newick_fns). Change if you want to run a subset of trees.
#' @runslow If FALSE, old, saved .Rdata results are loaded via load(). Default TRUE generates new .Rdata files and 
#' saves them via save()

source("run_bears_optim_on_multiple_trees2.R")

runslow = TRUE
if (runslow)
  {

  
	res = run_bears_optim_on_multiple_trees2(BioGeoBEARS_run_object, newick_fns,                                                        model_name="Glareosae_DEC_MT", geog_fns, resfns, run_or_collect="run", start_treenum=1, end_treenum=length(newick_fns), runslow=TRUE, plot_params=FALSE)

# Save res to "res_over_multiple_trees" 
	res_over_multiple_trees = res
	save(res_over_multiple_trees, file="res_over_multiple_trees.Rdata")
	} else {
	# Loads to "res_over_multiple_trees"
	load(file="res_over_multiple_trees.Rdata")
	} # END if (runslow)


# Now, collect the saved results from each individual run
res2 = run_bears_optim_on_multiple_trees2(BioGeoBEARS_run_object, newick_fns, model_name="Glareosae_DEC_MT", geog_fns, resfns, run_or_collect="collect", start_treenum=1, end_treenum=length(newick_fns), runslow=TRUE, plot_params=FALSE)
#res2


names(res2)

# This has the matrix of state probabilities at nodes
# for ALL of the trees you specified, in one huge matrix. 
dim(res2$state_probs_at_nodes_across_all_trees)
# The rightmost column is text, and contains the alphabetical, 
# comma-delimited list of tipnames (OTU names) descending from 
# that node or corner.
#
# These tipnames are used to identify "common" nodes between
# individual trees and the master tree. (See debates about
# the meaning of "common nodes", below.)
#
# The function summarize_stateprobs_on_master_tree(), below, is 
# what will go through this matrix, extract the probabilities
# for common nodes, and produce a matrix appropriate for
# plotting with standard functions.

# This matrix has probabilities for corners (branch bottoms), 
# instead of nodes.
dim(res2$state_probs_at_corners_across_all_trees)

# This has the averaged parameter values
res2$optim_results_mean


#################################################################
# To average the probabilities across all trees
#################################################################

# To plot with the standard BioGeoBEARS plot function you will need to 
# - run a standard analysis on the master tree, and 
# - then copy these average state probabilities into that res object 
#   (into the state probabilities at nodes, and state probabilities at branch bottoms)
# - then copy res2$optim_results_mean over the standard optim results


# Let's assume this tree is the master tree
best_trfn = "glareosae.tree"
tr = read.tree(best_trfn)
master_OTUs = tr$tip.label


stateprobs_list = summarize_stateprobs_on_master_tree(master_tree_fn=best_trfn, state_probs_at_nodes_across_all_trees=res2$state_probs_at_nodes_across_all_trees, state_probs_at_corners_across_all_trees=res2$state_probs_at_corners_across_all_trees, plotflag=FALSE)
stateprobs_list

#####################################################
# Get the single-tree master tree analysis
#####################################################
# Load the tipranges
tipranges = getranges_from_LagrangePHYLIP(lgdata_fn=geogfn)
tipranges

# Load master-tree analysis to "master_tree_res"
load(resfn)
resDEC_averaged = master_tree_res



#####################################################
# Input the averaged probabilities
#####################################################

# At nodes
resDEC_averaged$ML_marginal_prob_each_state_at_branch_top_AT_node = stateprobs_list$stateprobs_nodes_master_tr

# At corners
resDEC_averaged$ML_marginal_prob_each_state_at_branch_bottom_below_node = stateprobs_list$stateprobs_corners_master_tr

# Average parameter inferences
resDEC_averaged$optim_result = res2$optim_results_mean


#####################################################
# PLEASE NOTE: As you can see, 
# SOME NODES HAVE NO STATES OR PIE CHARTS. This is 
# because those nodes were not found in any of 
# the 3 trees sampled in example, above.
#
# In the probabilities matrix, these will be rows 
# that have "NA" instead of numbers.
#
# (Also see discussion of the ambiguity of what 
#  "same node" even means, in this context.)
#####################################################




#######################################################
# PDF plots
#######################################################
pdffn = "Plot_of_averages_on_master_tree_v1_DEC.pdf"
pdf(pdffn, width=6, height=6)

#######################################################
# Plot ancestral states - DEC
#######################################################
analysis_titletxt ="BioGeoBEARS DEC, averaged over 50 trees, plotted on master tree"

# Setup
results_object = resDEC_averaged
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.75, tipcex=0.3, statecex=0.3, splitcex=0.3, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.75, tipcex=0.3, statecex=0.5, splitcex=0.5, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)


dev.off()  # Turn off PDF
cmdstr = paste("open ", pdffn, sep="")
system(cmdstr) # Plot it




###############################################################
# Nick's advice/rant on the pseudo-Bayesian approach
###############################################################
# 
# After you've done all this, you'll have a plot of probabilities of states, 
# averaged over many trees. Congratulations.
# 
# However, I bet it will probably look pretty much like a plot of one run on the single ML or MCC tree.
# This is one of several reasons why I am skeptical of the popularity of "Bayes DIVA", "Bayes Lagrange", etc. 
# approaches. 
#
# SEVERAL POTENTIAL PROBLEMS WITH THE (PSEUDO) "BAYES" APPROACHES 
# TO HISTORICAL BIOGEOGRAPHY (Bayes DIVA, Bayes-Lagrange) 
#
# 1. First, they aren't really fully Bayesian -- more like pseudo-Bayesian.  
#
# 2. Second, the "Bayes" approach is sometimes done to mollify reviewers who heard somewhere 
# it was a good idea, probably because they think it helps "account for uncertainty." 
# However, The REALLY BIG uncertainty, is in things like:
#
# 2a. Model misspecification. For example, Bayes-DIVA is still giving DIVA results, and 
#     DIVA may be a poorly-fitting model on many datasets.  Running it over 2000 trees 
#     or whatever doesn't fix any of that, despite the elaborate and impressive-sounding
#     computational effort.
#
# 2b. Missing species (due to incomplete sampling, or extinction) will be missing in any
#     analysis, and this will not be fixed by running over many trees.
# 
# 2c. The "single most probable state" plots hide the uncertainty, because they just plot
#     the single-most probable estimated ancestral state at each node anyway. You need 
#     to do stochastic mapping to really feel, and characterize, the uncertainty.
#
# 3.  A third problem is that "averaging" "node probabilities" across trees is a weird 
#     operation. What is the definition of "same node" between two different trees?
#
# Consider these two clades, from two sampled trees:
# tree 1: ((A,B),C)
# tree 2: ((A,B),D)
#
# Is the common ancestor of (A,B) the "same" between the two trees?  The analysis above
# assumes "yes".
#
# (And, I think, all Bayes-DIVA/Bayes-Lagrange approaches assume this, except perhaps 
# the one I implemented in Wood, Matzke et al. (2013), Sys. Bio., which had more options.)
#
# Also, consider the "corner" at the bottom of the branch of the (A,B) node. Is *it* the same between
# the two trees?  The BioGeoBEARS program above assumes "yes".
#
# How about:
# tree 1: (((A,B),C),D)
# tree 2: (((A,C),B),D)
#
# Is the node that is common ancestor to ((A,B),C) "the same" as the node that is common ancestor to ((A,C),B)?
# Sort of, but not really. But, BioGeoBEARS and presumably the other programs above assume "yes".
#
# You can imagine getting "better matches" by requiring more constraints (same daughter tips, and same sister),
# but this will reduce matches between trees, and still doesn't really solve the problem, because unless
# trees agree perfectly in topology, there will always be mismatches of this sort.  And because biogeographical
# models have sudden cladogenetic transitions in addition to the continuous-time anagenetic ones, topological
# differences may have more significance than they would under typical, purely continuous-time models.
# 
#
# WHAT TO DO INSTEAD
#
# After thinking about it, my current conclusion is that the pseudo-Bayesian approaches
# need some serious critical thinking before being employed.
#
# Rather than running hundreds or thousands of trees and presenting dubious "average",
#
# * It would be better to spend time *thinking* about and discussing all of the sources
# of uncertainty, above, rather than spending days running a model on a bunch of (typically) very 
# similar trees that (typically) all say about the same thing.
#
# * It would be better to just run the analysis on a few different trees, and put those
#   plots in the paper.
#
# * If you are really concerned about some specific events, run a number of different trees, and then
#   do stochastic mapping on each one. The summary counts of events would be one way of summarizing
#   the uncertainty in event histories.  You could also identify some "common" nodes by some
#   explicitly-stated criterion, and count up the ranges inhabited and the counts of events at those nodes.
#
# The best argument for doing the pseudo-Bayesian approach is if you have a particular 
# node you are interested in, and the geography estimated at that node is strongly 
# affected by nearby uncertainty in the tree.  In that case,
# running a model over a bunch of trees can be useful.  
#
# I may be overly skeptical. But I have yet to see a case where the pseudo-Bayesian approach
# gave any real additional insight. It's only likely if your tree is very uncertain, but
# typically people get enough data to avoid that. Very often, I guess, people just want to
# be able to say, "we ran it on 100 trees and got the same result".  But really, if your
# tree is decently estimated, you actually knew that going in, from first principles.
#
# The pseudo-Bayesian approach might be useful if your dating is highly uncertain (as is common), 
# and you are doing a time-stratified analysis where uncertain nodes are near the time-strata.
# This would be the case I would focus on.
#


```


